---
title: "D5.2 – Uncertainty-aware Visual Analytic Components"
output:
  github_document:
    toc: true
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, magrittr, here, ggplot2, rstan, modelr, tidybayes, gganimate, splines, glue)
pacman::p_load(tsibble, hms, zoo)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

theme_set(
  theme_tidybayes()
)
knitr::opts_chunk$set(
  fig.width = 9,
  fig.height = 4.5,
  dev.args = list(type = "cairo")
)

set.seed(42)
```

## Load the data

Load the **current experiment** and the **past experiments**.
```{r, message = FALSE}
current_experiment <-
  read_csv(here::here("data", "interim", "running_experiment.csv")) %>%
  mutate(elapsedtime = hms(elapsedtime_s))
past_experiments <-
  read_csv(here::here("data", "interim", "past_experiments.csv")) %>%
  mutate(elapsedtime = hms(elapsedtime_s))
```

## Select the most similar experiments

Look for the fermentation experiments that have a smaller mean absolute error with our **experiment** until our **running_time**.
```{r, message = FALSE}
RULE <- "1:00:00" %>% as.hms() %>% as.numeric()
PCT_SELECTION <- .1

running_time <- max(current_experiment$elapsedtime)
experiment <- current_experiment$code_experimentation[1]

experiments_smooth <-
  past_experiments %>%
  filter(elapsedtime <= running_time) %>%
  bind_rows(current_experiment) %>%
  as_tsibble(
    index = elapsedtime,
    key = id(code_experimentation),
  ) %>%
  group_by(code_experimentation) %>%
  index_by(time = trunc_hms(elapsedtime, RULE)) %>%
  summarise(debitCO2_g = mean(debitCO2_g)) %>%
  fill_gaps(.full = TRUE) %>%
  mutate(debitCO2_g = na.approx(debitCO2_g, na.rm = FALSE, rule = 2)) %>%
  as_tibble()

experiments_error <- experiments_smooth %>%
  spread(code_experimentation, debitCO2_g) %>%
  select(-time) %>%
  mutate_all(`-`, .[[experiment]]) %>%
  mutate_all(abs) %>%
  summarise_all(mean) %>%
  gather(code_experimentation, MAE) %>%
  arrange(MAE) %>%
  filter(code_experimentation != experiment) %>%
  filter(MAE < quantile(MAE, PCT_SELECTION))


selected_experiments <- experiments_error %>% pull(code_experimentation)
```

Plot the selected experiments (the 10% with smaller error).
```{r}
experiments_smooth %>%
  mutate(exeriments = ifelse(code_experimentation %in% selected_experiments, "selected", "non selected")) %>%
  mutate(exeriments = ifelse(code_experimentation == experiment, "running experiment", exeriments)) %>%
  ggplot(aes(time, debitCO2_g)) +
  geom_line(aes(group = code_experimentation, color = exeriments, alpha = exeriments)) +
  scale_color_manual(values = c("grey", "navy", "skyblue")) +
  scale_alpha_manual(values = c(0.3, 1, 0.7)) +
  labs(x = "Elapsed Time", y = "debitCO2 (g)")
```

## Bayesian Model

Smooth the past experiments by calculating the mean over each hour. 

```{r}
RULE_OBS <- "1:00:00" %>% as.hms() %>% as.numeric()

obs <- past_experiments %>%
  filter(code_experimentation %in% selected_experiments) %>%
  rename(elapsedtime_reading = elapsedtime) %>%
  as_tsibble(
    index = elapsedtime_reading,
    key = id(code_experimentation),
  ) %>%
  group_by(code_experimentation) %>%
  index_by(elapsedtime = trunc_hms(elapsedtime_reading, RULE_OBS)) %>%
  summarise(debitCO2_g = mean(debitCO2_g)) %>%
  as_tibble() %>%
  filter(elapsedtime <= quantile(elapsedtime, .9)) %>%
  arrange(elapsedtime)
```

Calculate the data needed by the model, compile the .stan file and sample from it (or load if already sampled).

```{r, message = FALSE, error = FALSE}
X <- obs$elapsedtime
Y <- obs$debitCO2_g
num_data <- length(X)

num_knots <- 15
knot_list <- quantile(obs$elapsedtime, probs = seq(0, 1, length.out = num_knots))
B <- bs(obs$elapsedtime,
  knots = knot_list[-c(1, num_knots)],
  degree = 3, intercept = TRUE
)
num_basis <- ncol(B)

data <- list(X = X, Y = Y, num_data = num_data, num_basis = num_basis)
model_file <- glue("m_{experiment}_{running_time}s.rds")

if (file.exists(here::here("data", "interim", model_file))) {
  fit <- readRDS(here("data", "interim", model_file))
} else {
  sm <- stan_model(here::here("src", "models", "fermentation.stan"))
  fit <- sampling(sm, data = data, iter = 500, control = list(adapt_delta = 0.95))
  saveRDS(fit, here::here("data", "interim", model_file))
}
```

Draw predictions from the model samples.
```{r}
num_points <- 100
elapsedtime_quantiles <- quantile(obs$elapsedtime, probs = seq(0, 1, length.out = num_points - 1)) %>% unname()
elapsedtime_seq <- sort(c(as.hms(running_time), elapsedtime_quantiles))

new_B <- bs(elapsedtime_seq,
  knots = knot_list[-c(1, num_knots)],
  degree = 3, intercept = TRUE
)

post_samples <- fit %>%
  spread_draws(w[bi], w_sigma[bi]) %>%
  group_by(.chain, .iteration, .draw) %>%
  do(.value = new_B %*% as.vector(t(.$w)), sigma = new_B %*% as.vector(t(.$w_sigma))) %>%
  mutate(elapsedtime = list(elapsedtime_seq)) %>%
  unnest() %>%
  mutate(.prediction = rnorm(n(), .value, sigma)) %>%
  select(.draw, elapsedtime, .prediction, .value, )

post_samples_with_last_obs <-
  current_experiment %>%
  slice(n()) %>%
  mutate(.draw = list(1:max(post_samples$.draw))) %>%
  unnest() %>%
  mutate(.prediction = debitCO2_g, .value = debitCO2_g) %>%
  select(.draw, elapsedtime, .prediction, .value, ) %>%
  bind_rows(post_samples %>% filter(elapsedtime > running_time)) %>%
  filter(elapsedtime >= running_time) %>%
  arrange(.draw, elapsedtime)
```

## Uncertainty Visualizations
```{r}
labs <- labs(
  title = glue("Fermentation of Experiment {experiment} ({as.hms(running_time)})"),
  x = "Elapsed Time", y = "debitCO2 (g)"
)
```


### Spaghetti 

```{r}
time_limit <- elapsedtime_seq[match(TRUE, elapsedtime_seq >= running_time)]

prediction_color <- "#3573b9"
past_color <- "gray1"
num_samples <- 100
samples <- post_samples %>% sample_draws(num_samples)

spaghetti_plot <-
  current_experiment %>%
  ggplot(aes(x = elapsedtime, y = debitCO2_g)) +
  geom_line(aes(y = .prediction, group = .draw),
    color = prediction_color, alpha = 1 / 20, size = .75,
    data = samples %>% filter(elapsedtime >= time_limit)
  ) +
  geom_line(aes(y = .prediction, group = .draw),
    color = past_color, alpha = 1 / 40, size = .75,
    data = samples %>% filter(elapsedtime <= time_limit)
  ) +
  geom_line(size = 1.25, color = "gray20") +
  labs

spaghetti_plot
```

Load and show the real evolution of the experiment fermentation.

```{r message = FALSE}
experiment_future <-
  read_csv(here::here("data", "interim", "experiment_future.csv")) %>%
  mutate(elapsedtime = hms(elapsedtime_s))

future_line <- geom_line(size = 1.25, color = "#ff7f0e", data = experiment_future)

spaghetti_plot + future_line
```


### Uncertainty bands
```{r}
current_experiment %>%
  ggplot(aes(x = elapsedtime, y = debitCO2_g)) +
  stat_lineribbon(aes(y = .prediction),
    fill = adjustcolor(prediction_color, alpha.f = .2), color = prediction_color,
    data = post_samples %>% filter(elapsedtime >= time_limit), .width = .95
  ) +
  stat_lineribbon(aes(y = .prediction),
    fill = adjustcolor(past_color, alpha.f = .2), color = NA,
    data = post_samples %>% filter(elapsedtime <= time_limit), .width = .95
  ) +
  geom_line(size = 1.25, color = "gray20") +
  labs
```


```{r}
bands_plot <-
  current_experiment %>%
  ggplot(aes(x = elapsedtime, y = debitCO2_g)) +
  stat_lineribbon(aes(y = .prediction),
    fill = prediction_color, alpha = 1 / 5,
    data = post_samples %>% filter(elapsedtime >= time_limit)
  ) +
  stat_lineribbon(aes(y = .prediction),
    fill = past_color, color = NA, alpha = 1 / 5,
    data = post_samples %>% filter(elapsedtime <= time_limit)
  ) +
  geom_line(size = 1.25, color = "gray20") +
  labs

bands_plot
```

```{r}
bands_plot + future_line
```


### Gradient
```{r}
n_bands <- 40

gradient_plot <-
  current_experiment %>%
  ggplot(aes(x = elapsedtime, y = debitCO2_g)) +
  stat_lineribbon(aes(y = .prediction),
    fill = prediction_color, color = NA, alpha = 1 / n_bands, .width = ppoints(n_bands),
    data = post_samples %>% filter(elapsedtime >= time_limit)
  ) +
  stat_lineribbon(aes(y = .prediction),
    fill = past_color, color = NA, alpha = 1 / n_bands, .width = ppoints(n_bands),
    data = post_samples %>% filter(elapsedtime <= time_limit)
  ) +
  geom_line(size = 1.25, color = "gray20") +
  labs


gradient_plot
```

```{r}
gradient_plot + future_line
```

### Hypothetical Outcome Plots (HOPs)

```{r message = FALSE}
samples_hop <- post_samples_with_last_obs %>% filter(elapsedtime >= running_time) %>% sample_draws(num_samples)

hop <- current_experiment %>%
  ggplot(aes(x = elapsedtime, y = debitCO2_g)) +
  geom_line(aes(y = .prediction, group = .draw),
    color = prediction_color, size = .75,
    data = samples_hop
  ) +
  geom_line(size = 1.25, color = "gray20") +
  labs +
  transition_states(.draw, 0, 1)

animate(hop,
  nframes = num_samples, fps = 2.5, res = 100,
  width = 900, height = 450, type = "cairo"
)
```

```{r}
hop_spaghetti <- hop +
  shadow_mark(past = TRUE, future = TRUE, color = "gray1", alpha = 1 / 40)

animate(hop_spaghetti,
  nframes = num_samples, fps = 2.5, res = 100,
  width = 900, height = 450, type = "cairo"
)
```

```{r message = FALSE}
hop$layers <- c(future_line, hop$layers)

animate(hop,
  nframes = num_samples, fps = 2.5, res = 100,
  width = 900, height = 450, type = "cairo"
)
```

## Session Info

```{r echo = FALSE}
sessionInfo()
```

